---
title: "Varimax CLT paper - numerical example - undirected SBM"
author:
- Joshua Cape (jrcape@wisc.edu)
output: html_notebook
---

## Preamble
```{r, message=FALSE, warning=FALSE}
rm(list = ls())
source("VCLT_preamble.R")
source("VCLT_sbms_model_RoheSPCA.R")
source("VCLT_sbms_undirected_covmtx.R")
```

### Helper function for converting R matrix output to LaTeX code
```{r}
# https://www.r-bloggers.com/2020/08/matrix-to-latex/
array_to_LaTeX <- function(arr){
  rows <- apply(arr, MARGIN=1, paste, collapse = " & ")
  matrix_string <- paste(rows, collapse = " \\\\ ")
  return(cat(paste("\\begin{bmatrix}", matrix_string, "\\end{bmatrix}")))
  }
```

## Show theoretical covariance matrices for sourced SBM model
```{r}
if(rho == 1)
  vmx.cov.theory <- lapply(1:K, function(xxx)
    cov_mtx_vmx(as.vector(T.mtx %*% diag(K)[xxx,])))
if(rho < 1)
  vmx.cov.theory <- lapply(1:K, function(xxx)
    cov_mtx_vmx_sparse(as.vector(T.mtx %*% diag(K)[xxx,])))
print(vmx.cov.theory)
```

### Print LaTeX code for a theoretical covariance matrix (numerically rounded)
```{r}
array_to_LaTeX(signif(vmx.cov.theory[[1]], 3))
```

### Print LaTeX code for a theoretical covariance matrix (numerically rounded)
```{r}
array_to_LaTeX(signif(vmx.cov.theory[[2]], 3))
```

## Single simulation trial - generating a single SBM graph
```{r}

  set.seed(1234)
  source("VCLT_sbms_undirected.R")

  U.hat <-
    as.matrix(eigs(as_adjacency_matrix(my.graph), d, which="LM")$vectors)
  
      nonzero.row.index.Uhat <-
        min(which(sapply(1:K, function(xxx) min(abs(U.hat[xxx,])) > n^(-2) ) == TRUE))
      # << Safety check: verifies that no row of U.hat is numerically zero / truncated eigendecomposition failed.
  
          U.hat <- U.hat %*% diag(sign(U.hat[nonzero.row.index.Uhat,]), nrow=d)
  
  Z.hat <- sqrt(n) * varimax(U.hat, normalize = FALSE)$loadings
  
        # Align raw embedding via permutation/signflip matrix (P^T from theory)
        m1 <- diag(1/sqrt(pi.vec));
        m2 <- Z.hat[cumsum(block.sizes),]; #<<< uses candidate vectors from each block
        orien.mtx <- round(svd(t(m2) %*% m1)$u %*% t(svd(t(m2) %*% m1)$v));
        Z.hat <- Z.hat %*% orien.mtx

# NB: normalized embedding vectors          
  vmx.cov.raw.est <-
    lapply(1:K, function(xxx) Mclust(sqrt(n * rho) * (Z.hat[memb.index==xxx,] - Z.mtx[memb.index==xxx,] %*% diag(1/sqrt(pi.vec))), G=1, verbose = FALSE, modelNames="VVV")$parameters$variance$sigma[,,1]); vmx.cov.raw.est

```

### Show simulation output as LaTeX source code
```{r}
array_to_LaTeX(signif(vmx.cov.raw.est[[1]], 3))
```

### Show simulation output as LaTeX source code
```{r}
array_to_LaTeX(signif(vmx.cov.raw.est[[2]], 3))
```


## Multiple graph simulation
```{r}
set.seed(1234)

num.graphs <- 100
data.cov.errors <- list()
data.Zi.stand.estimates <- list()

#--------------------------------------------------------
for(rep in 1:num.graphs){
  source("VCLT_sbms_undirected.R")

  U.hat <- as.matrix(eigs(as_adjacency_matrix(my.graph), d, which="LM")$vectors)
    nonzero.row.index.Uhat <- min(which(sapply(1:K, function(xxx) min(abs(U.hat[xxx,])) > n^(-2)) == TRUE))
      U.hat <- U.hat %*% diag(sign(U.hat[nonzero.row.index.Uhat,]), nrow=d)
  
  Z.hat <- sqrt(n) * varimax(U.hat, normalize = FALSE)$loadings
  
        m1 <- diag(1/sqrt(pi.vec));
        m2 <- Z.hat[cumsum(block.sizes),];
        orien.mtx <- round(svd(t(m2) %*% m1)$u %*% t(svd(t(m2) %*% m1)$v));
        Z.hat <- Z.hat %*% orien.mtx

  vmx.cov.raw.est <- lapply(1:K, function(xxx) Mclust(sqrt(n * rho) * (Z.hat[memb.index==xxx,] -  Z.mtx[memb.index==xxx,] %*% diag(1/sqrt(pi.vec))), G=1, verbose = FALSE, modelNames="VVV")$parameters$variance$sigma[,,1])
  
#---------- output below ---------------#
  
  data.cov.errors[[rep]] <-
    sapply(1:K, function(xxx) norm(vmx.cov.raw.est[[xxx]] - vmx.cov.theory[[xxx]], "F")/norm(vmx.cov.theory[[xxx]], "F"))

  data.Zi.stand.estimates[[rep]] <-
    sapply(1:K, function(xxx) (sqrt(n)*(colMeans(Z.hat[memb.index==xxx,]) - diag(1/sqrt(pi.vec))[xxx,]))[xxx])
    
  }
```


### Store simulated data
```{r}
#  write.table(t(mapply(c,data.cov.errors)),
#   file="paper_VCLT_data_SBM_cov_error_estimates.txt")

#  write.table(t(mapply(c,data.Zi.stand.estimates)),
#    file="paper_VCLT_data_SBM_Zi_stand_estimates.txt")
```


## Summarize numerical output in table form
```{r}
signif(apply(t(mapply(c,data.cov.errors)), 2, mean), 4)*100 # mean relative Frobenius norm error
signif(apply(t(mapply(c,data.cov.errors)), 2, sd)/sqrt(num.graphs), 4)*100 # sample standard error
```


### Run Shapiro-Wilk test for normality
```{r}
for(i in 1:K) print(shapiro.test(t(mapply(c,data.Zi.normality))[,i]))
```

